---
title: "Functional Analysis"
author: "Liz Mallott"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Loading required libraries

The following libraries are required for this analysis:
tidyverse (v.1.3.0)
vegan (v.2.5-6)
pairwiseAdonis (v.0.3, https://github.com/pmartinezarbizu/pairwiseAdonis)

```{r load_libraries}
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(ggplot2)
library(cowplot)
library(glmmTMB)
library(multcomp)
library(car)
library(fdrtool)
```

## Importing data

Bray-Curtis and Jaccard distance matrices were created in QIIME2 for both gene family and pathway abundances so we could easily rarefy the data and remove low-abundance samples. So, we are importing those matrices here, after ensuring they 

```{r import_data}
bray_gf = as.dist(read.table("distance-matrix-bray-gf.tsv", header = T))
bray_gf_full = as.dist(read.table("distance-matrix-bray-gf-full.tsv", header = T))
jaccard_gf = as.dist(read.table("distance-matrix-jaccard-gf.tsv", header = T))
jaccard_gf_full = as.dist(read.table("distance-matrix-jaccard-gf-full.tsv", header = T))
bray_pa = as.dist(read.table("distance-matrix-bray-pa.tsv", header = T))
bray_pa_full = as.dist(read.table("distance-matrix-bray-pa-full.tsv", header = T))
jaccard_pa = as.dist(read.table("distance-matrix-jaccard-pa.tsv", header = T))
jaccard_pa_full = as.dist(read.table("distance-matrix-jaccard-pa-full.tsv", header = T))
bray_cog_full = as.dist(read.table("distance-matrix-bray-cog-full.tsv", header = T))
jaccard_cog_full = as.dist(read.table("distance-matrix-jaccard-cog-full.tsv", header = T))

metadata_gf = read.csv("chioma_Sept20_metadata_gf.csv", header = T)
metadata_gf_full = read.csv("chioma_Sept20_metadata_gf_full.csv", header = T)
metadata_pa = read.csv("chioma_Sept20_metadata_pa.csv", header = T)
metadata_pa_full = read.csv("chioma_Sept20_metadata_pa_full.csv", header = T)
metadata_cog_full = read.csv("chioma_Sept20_metadata_cog_full.csv", header = T)
```

## PERMANOVAs

A PERMANOVA is performed to assess the influence of both floor (7th vs. 8th vs. germ-free) and treatment (bleomycin vs. saline). 

We will first look at pathway abundance taking abundance into account (Bray-Curtis metric), running matrices rarefied at two different depths.

```{r permanova_bray_pa}
adonis2(bray_pa ~ Floor + Treatment, data = metadata_pa, 
        by = "margin", permutations = 4999)
adonis2(bray_pa_full ~ Floor + Treatment, data = metadata_pa_full, 
        by = "margin", permutations = 4999)
```

And then at presence-absence (Jaccard metric).

```{r permanova_jaccard_pa}
adonis2(jaccard_pa ~ Floor + Treatment, data = metadata_pa, 
        by = "margin", permutations = 4999)
adonis2(jaccard_pa_full ~ Floor + Treatment, data = metadata_pa_full, 
        by = "margin", permutations = 4999)
adonis2(jaccard_gf_full ~ Floor + Treatment, data = metadata_gf_full, 
        by = "margin", permutations = 4999)
```

Pairwise comparisons between floors for the dataset that still includes germ free treatments are also performed.

```{r pairwise_bray_pa}
pairwise.adonis(bray_pa_full, factors = metadata_pa_full$Floor, perm = 4999, p.adjust.m = "holm")
```

```{r pairwise_jaccard_pa}
pairwise.adonis(jaccard_pa_full, factors = metadata_pa_full$Floor, perm = 4999, p.adjust.m = "holm")
```

We will then look at gene family abundance taking abundance into account (Bray-Curtis metric), running matrices rarefied at two different depths.

```{r permanova_bray_gf}
adonis2(bray_gf ~ Floor + Treatment, data=metadata_gf, 
        by = "margin", permutations = 4999)
adonis2(bray_gf_full ~ Floor + Treatment, data=metadata_gf_full, 
        by = "margin", permutations = 4999)
```

And then at presence-absence (Jaccard metric).

```{r permanova_jaccard_gf}
adonis2(jaccard_gf ~ Floor + Treatment, data=metadata_gf, 
        by = "margin", permutations = 4999)
adonis2(jaccard_gf_full ~ Floor + Treatment, data=metadata_gf_full, 
        by = "margin", permutations = 4999)
```

Pairwise comparisons between floors are also performed.

```{r pairwise_bray_gf}
pairwise.adonis(bray_gf_full, factors = metadata_gf_full$Floor, perm = 4999, p.adjust.m = "holm")
```
We will then look at clusters of orthologous groups (COGs) taking abundance into account (Bray-Curtis metric).
```{r permanova_bray_cog}
adonis2(bray_cog_full ~ Floor + Treatment, data=metadata_cog_full, 
        by = "margin", permutations = 4999)
```

And then at presence-absence (Jaccard metric).

```{r permanova_jaccard_cog}
adonis2(jaccard_cog_full ~ Floor + Treatment, data=metadata_cog_full, 
        by = "margin", permutations = 4999)
```

Pairwise comparisons between floors are also performed.

```{r pairwise_bray_cog}
pairwise.adonis(bray_cog_full, factors = metadata_cog_full$Floor, perm = 4999, p.adjust.m = "holm")
```

```{r pairwise_jaccard_cog}
pairwise.adonis(jaccard_cog_full, factors = metadata_cog_full$Floor, perm = 4999, p.adjust.m = "holm")
```


## NMDS plots

We constructed a non-linear multidimensional scaling plot for the Bray-Curtis metric using the larger datasets for both pathway abundance and gene families.

```{r nmds}
mds_bray_pa<-metaMDS(bray_pa_full, k=2, trymax=500)
mds_bray_pa_points<-mds_bray_pa$points
mds_bray_pa_points2<-merge(x=mds_bray_pa_points, y = metadata_pa_full, 
                                  by.x = "row.names", by.y = "sample.id")
braypa <- ggplot(mds_bray_pa_points2, aes(x = MDS1, y = MDS2, 
                                                 color = Floor, shape = Treatment)) +
  geom_point(size=3) + scale_color_brewer(palette = 'Set1') +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'), 
        legend.key=element_blank()) + 
  theme(axis.title.x=element_text(size=rel(2)), 
        axis.title.y=element_text(size=rel(2)),
        plot.title = element_text(size=rel(3)),
        legend.title = element_text(size=rel(2)),
        legend.text = element_text(size = rel(1.8))) + 
  ggtitle("Pathway abundance")
braypa

mds_bray_gf<-metaMDS(bray_gf_full, k=2, trymax=500)
mds_bray_gf_points<-mds_bray_gf$points
mds_bray_gf_points2<-merge(x=mds_bray_gf_points, y = metadata_gf_full, 
                                  by.x = "row.names", by.y = "sample.id")
braygf <- ggplot(mds_bray_gf_points2, aes(x = MDS1, y = MDS2, 
                                                 color = Floor, shape = Treatment)) +
  geom_point(size=3) + scale_color_brewer(palette = 'Set1') +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'), 
        legend.key=element_blank()) + 
  theme(axis.title.x=element_text(size=rel(2)), 
        axis.title.y=element_text(size=rel(2)),
        plot.title = element_text(size=rel(3)),
        legend.title = element_text(size=rel(2)),
        legend.text = element_text(size = rel(1.8))) + 
  ggtitle("Gene families")
braygf

tiff(file = "nmds_plot_combined_shotgun.tif", res = 300, width = 18, height = 7, units="in")
legend1 = get_legend(braygf + theme(legend.box.margin = margin(0, 0, 0, 12)))
plot_grid(braygf + theme(legend.position = "none"), 
          braypa + theme(legend.position = "none"), legend1, 
          nrow = 1, ncol = 3, rel_widths = c(1.5, 1.5, 0.75), align = "h", 
          axis = "t")
dev.off()
```
PDF plots are included below
```{r bray_pdf}
pdf(file="nmds_plot_combined_shotgun.pdf", width=18, height=7)
legend1 = get_legend(braygf + theme(legend.box.margin = margin(0, 0, 0, 12)))
plot_grid(braygf + theme(legend.position = "none"), 
          braypa + theme(legend.position = "none"), legend1, 
          nrow = 1, ncol = 3, rel_widths = c(1.5, 1.5, 0.75), align = "h", 
          axis = "t")
dev.off()
```
## Individual pathway differences

We ran a generalized linear model with a negative binomial distribution to assess the influence of treatment and floor on the relative abundance of all pathways detected.

```{r path_comp}
paths = read_tsv("chioma_pathabundance_relab_unstratified.tsv")
paths_t = paths[, colSums(paths != 0) > 0] %>% pivot_longer(-Pathway) %>% 
  pivot_wider(names_from=Pathway, values_from=value)

metadata = read_tsv("chioma_Sept20_metadata_function.txt", col_types = "cccfffddfffcf")

paths_meta = inner_join(metadata, paths_t, by = "name") %>% filter(Treatment == "Saline" | Treatment == "Bleomycin") 


paths_meta = as.data.frame(paths_meta)

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

floor_estimatematrix = mat.or.vec(301,2)
floor_pvaluematrix = mat.or.vec(301,2)
treatment_estimatematrix = mat.or.vec(301,2)
treatment_pvaluematrix = mat.or.vec(301,2)

for(i in 14:314) {
  variable = paths_meta[,i]
  b = try(glmmTMB(variable ~ Floor + Treatment, 
                  data = paths_meta, family = nbinom2))
  anova = Anova(b)
  floor_estimatematrix[i-13,2] = anova[1,1]
  floor_pvaluematrix[i-13,2] = anova[1,3]
  treatment_estimatematrix[i-13,2] = anova[2,1]
  treatment_pvaluematrix[i-13,2] = anova[2,3]
  floor_estimatematrix[i-13,1] = names(paths_meta)[i]
  floor_pvaluematrix[i-13,1] = names(paths_meta)[i]
  treatment_estimatematrix[i-13,1] = names(paths_meta)[i]
  treatment_pvaluematrix[i-13,1] = names(paths_meta)[i]
}

floor = bind_cols(as.data.frame(floor_estimatematrix[,1:2]), 
                     as.data.frame(floor_pvaluematrix[,2]))
write.csv(floor, "floor_pathways.csv")
floor_nona = filter(floor, floor[,3] != "NaN") 
floor_nona[,3] = as.numeric(as.character(floor_nona[,3]))
floor_corrected = bind_cols(floor_nona, 
                               as.data.frame(fdrtool(floor_nona[,3], 
                                                     statistic = "pvalue", plot = F)))
write.csv(floor_corrected, "floor_corrected_pathways.csv")

treatment = bind_cols(as.data.frame(treatment_estimatematrix[,1:2]), 
                      as.data.frame(treatment_pvaluematrix[,2]))
write.csv(treatment, "treatment_pathways.csv")
treatment_nona = filter(treatment, treatment[,3] != "NaN") 
treatment_nona[,3] = as.numeric(as.character(treatment_nona[,3]))
treatment_corrected = bind_cols(treatment_nona, 
                                as.data.frame(fdrtool(treatment_nona[,3], 
                                                      statistic = "pvalue", plot = F)))
write.csv(treatment_corrected, "treatment_corrected_pathways.csv")

```

## Individual COG differences

We ran a generalized linear model with a negative binomial distribution to assess the influence of treatment and floor on the  abundance of all COGs detected.

```{r cog_comp}
cogs = read_tsv("chioma_genefamilies_cog_cpm_unstratified.tsv")
cogs_t = cogs[, colSums(cogs != 0) > 0] %>% pivot_longer(-GeneFamily) %>% 
  pivot_wider(names_from=GeneFamily, values_from=value)

metadata = read_tsv("chioma_Sept20_metadata_function.txt", col_types = "cccfffddfffcf")

cogs_meta = inner_join(metadata, cogs_t, by = "name") %>% filter(Treatment == "Saline" | Treatment == "Bleomycin") 


cogs_meta = as.data.frame(cogs_meta)

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

floor_estimatematrix = mat.or.vec(12131,2)
floor_pvaluematrix = mat.or.vec(12131,2)
treatment_estimatematrix = mat.or.vec(12131,2)
treatment_pvaluematrix = mat.or.vec(12131,2)

for(i in 14:12144) {
  variable = cogs_meta[,i]
  b = try(glmmTMB(variable ~ Floor + Treatment, 
                  data = cogs_meta, family = nbinom2))
  anova = Anova(b)
  floor_estimatematrix[i-13,2] = anova[1,1]
  floor_pvaluematrix[i-13,2] = anova[1,3]
  treatment_estimatematrix[i-13,2] = anova[2,1]
  treatment_pvaluematrix[i-13,2] = anova[2,3]
  floor_estimatematrix[i-13,1] = names(cogs_meta)[i]
  floor_pvaluematrix[i-13,1] = names(cogs_meta)[i]
  treatment_estimatematrix[i-13,1] = names(cogs_meta)[i]
  treatment_pvaluematrix[i-13,1] = names(cogs_meta)[i]
}

floor = bind_cols(as.data.frame(floor_estimatematrix[,1:2]), 
                     as.data.frame(floor_pvaluematrix[,2]))
write.csv(floor, "floor_cogs.csv")
floor_nona = filter(floor, floor[,3] != "NaN") 
floor_nona[,3] = as.numeric(as.character(floor_nona[,3]))
floor_corrected = bind_cols(floor_nona, 
                               as.data.frame(fdrtool(floor_nona[,3], 
                                                     statistic = "pvalue", plot = F)))
write.csv(floor_corrected, "floor_corrected_cogs.csv")

floor_corrected_sig = floor_corrected %>% 
  filter(qval <= 0.05)
write.csv(floor_corrected_sig, "floor_corrected_sig_cogs.csv")

treatment = bind_cols(as.data.frame(treatment_estimatematrix[,1:2]), 
                      as.data.frame(treatment_pvaluematrix[,2]))
write.csv(treatment, "treatment_cogs.csv")
treatment_nona = filter(treatment, treatment[,3] != "NaN") 
treatment_nona[,3] = as.numeric(as.character(treatment_nona[,3]))
treatment_corrected = bind_cols(treatment_nona, 
                                as.data.frame(fdrtool(treatment_nona[,3], 
                                                      statistic = "pvalue", plot = F)))
write.csv(treatment_corrected, "treatment_corrected_cogs.csv")

```
