---
title: "Taxonomic Analysis"
author: "Liz Mallott"
date: "9/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading required libraries

The following libraries are required for this analysis:
tidyverse (v.1.3.0)
vegan (v.2.5-6)
pairwiseAdonis (v.0.3, https://github.com/pmartinezarbizu/pairwiseAdonis)
cowplot (v.1.1.0)
glmmTMB (v.1.0.2.1)
multcomp (v.1.4-14)
car (v.3.0-9)
fdrtool (v.1.2.15)

```{r load_libraries}
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(cowplot)
library(glmmTMB)
library(multcomp)
library(car)
library(fdrtool)
library(ggpubr)
```

## Importing data

The species-level merged relative abundance table from MetaPhlan and metadata files are loaded. Samples with no reads mapping to bacteria (9 germ-free samples) and all control samples are removed and the relative abundance table is transposed for ease of use in vegan.

```{r import_data}
species = read_tsv("merged_abundance_table_species.txt")
metadata = read_tsv("chioma_Sept20_metadata.txt", col_types = "cccfffddfffcf")
species_t = species[, colSums(species != 0) > 0] %>% pivot_longer(-body_site) %>% 
  pivot_wider(names_from=body_site, values_from=value)
species_meta = inner_join(metadata, species_t) %>% filter(Treatment == "Saline" | Treatment == "Bleomycin") 
```

## Calculating distance matrices

Bray-Curtis and Jaccard distance matrices are calculated in vegan.

```{r distances}
bray = vegdist(species_meta[,14:55], method = "bray")
jaccard = vegdist(species_meta[14:55], method = "jaccard")
```

## PERMANOVAs

A PERMANOVA is performed to assess the influence of both floor (7th vs. 8th vs. germ-free) and treatment (bleomycin vs. saline). 

We will first look at composition taking species abundance into account (Bray-Curtis metric).

```{r permanova_bray}
adonis2(bray ~ Floor + Treatment, data=species_meta, 
        by = "margin", permutations = 4999)
```

And then at presence-absence (Jaccard metric).

```{r permanova_jaccard}
adonis2(jaccard ~ Floor + Treatment, data=species_meta, 
        by = "margin", permutations = 4999)
```

Pairwise comparisons between floors are also performed.

```{r pairwise_bray}
pairwise.adonis(bray, factors = species_meta$Floor, perm = 4999, p.adjust.m = "holm")
```

```{r pairwise_jaccard}
pairwise.adonis(jaccard, factors = species_meta$Floor, perm = 4999, p.adjust.m = "holm")
```

## NMDS plots

We constructed a non-linear multidimensional scaling plot for the Bray-Curtis metric.

```{r nmds_bray}
mds_otus_bray<-metaMDS(bray, k=2, trymax=500)
mds_otus_bray_points<-mds_otus_bray$points
mds_otus_bray_points2<-merge(x=mds_otus_bray_points, y = species_meta, 
                                  by.x = "row.names", by.y = "row.names")
tiff(file="nmds_plot_bray.tif", res=300, width=8, height=6, units="in")
bray <- ggplot(mds_otus_bray_points2, aes(x = MDS1, y = MDS2, 
                                                 color = Floor, shape = Treatment)) +
  geom_point(size=3) + scale_color_brewer(palette = 'Set1') +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'), 
        legend.key=element_blank()) + 
  theme(axis.title.x=element_text(size=rel(2)), 
        axis.title.y=element_text(size=rel(2)),
        plot.title = element_text(size=rel(3)),
        legend.title = element_text(size=rel(2)),
        legend.text = element_text(size = rel(1.8))) + 
  ggtitle("Bray-Curtis")
bray
dev.off()
bray
```
PDF plots are included below
```{r bray_pdf}
pdf(file="nmds_plot_bray.pdf", width=8, height=6)
bray
dev.off()
```

## Individual taxonomic differences

We ran a generalized linear model with a negative binomial distribution to assess the influence of treatment and floor on the relative abundance of all species detected.

```{r taxa_comp}
species_meta_comp = as.data.frame(species_meta)

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

floor_estimatematrix = mat.or.vec(42,2)
floor_pvaluematrix = mat.or.vec(42,2)
treatment_estimatematrix = mat.or.vec(42,2)
treatment_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_comp[,i]
  b = try(glmmTMB(variable ~ Floor + Treatment, 
                  data = species_meta_comp, family = nbinom2))
  anova = Anova(b)
  floor_estimatematrix[i-13,2] = anova[1,1]
  floor_pvaluematrix[i-13,2] = anova[1,3]
  treatment_estimatematrix[i-13,2] = anova[2,1]
  treatment_pvaluematrix[i-13,2] = anova[2,3]
  floor_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  floor_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
}

floor = bind_cols(as.data.frame(floor_estimatematrix[,1:2]), 
                     as.data.frame(floor_pvaluematrix[,2]))
write.csv(floor, "floor.csv")
floor_nona = filter(floor, floor[,3] != "NaN") 
floor_nona[,3] = as.numeric(as.character(floor_nona[,3]))
floor_corrected = bind_cols(floor_nona, 
                               as.data.frame(fdrtool(floor_nona[,3], 
                                                     statistic = "pvalue", plot = F)))
write.csv(floor_corrected, "floor_corrected.csv")

treatment = bind_cols(as.data.frame(treatment_estimatematrix[,1:2]), 
                      as.data.frame(treatment_pvaluematrix[,2]))
write.csv(treatment, "treatment.csv")
treatment_nona = filter(treatment, treatment[,3] != "NaN") 
treatment_nona[,3] = as.numeric(as.character(treatment_nona[,3]))
treatment_corrected = bind_cols(treatment_nona, 
                                as.data.frame(fdrtool(treatment_nona[,3], 
                                                      statistic = "pvalue", plot = F)))
write.csv(treatment_corrected, "treatment_corrected.csv")

a = glmmTMB(Parasutterella_excrementihominis ~ Floor + Treatment, 
            data = species_meta_comp, family = nbinom2)
summary(a)
Anova(a)
summary(glht(a,linfct=mcp(Floor="Tukey")))
```

## Single Taxa Graphs

One taxa, Parasutterella excrementihominis, was more abundant in mice on the non-infectious floor compared with both the infectious and germ-free floors.

```{r graphs}
para = ggboxplot(species_meta_comp, x = "Floor", 
                    y = "Parasutterella_excrementihominis", color = "Floor", 
                    palette = "Set1", add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Relative Abundance", title = "Parasutterella excrementihominis") 
para = ggpar(para, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title")
para

tiff(file="Floor_taxa.tif", res=300, width=4, height=3, units="in")
para
dev.off()
```

And PDF output
```{r graphs_pdf}
pdf("Floor_taxa.pdf", width = 4, height = 3)
para
dev.off()
```
