---
title: "Taxonomic Analysis"
author: "Liz Mallott"
date: "9/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading required libraries

The following libraries are required for this analysis:
tidyverse (v.1.3.0)
vegan (v.2.5-6)
pairwiseAdonis (v.0.3, https://github.com/pmartinezarbizu/pairwiseAdonis)
cowplot (v.1.1.0)
glmmTMB (v.1.0.2.1)
multcomp (v.1.4-14)
car (v.3.0-9)
fdrtool (v.1.2.15)

```{r load_libraries}
library(tidyverse)
library(vegan)
library(pairwiseAdonis)
library(cowplot)
library(glmmTMB)
library(multcomp)
library(car)
library(fdrtool)
library(ggpubr)
```

## Importing data

The species-level merged relative abundance table from MetaPhlan and metadata files are loaded. All germ-free and all control samples are removed and the relative abundance table is transposed for ease of use in vegan.

```{r import_data}
species = read_tsv("merged_abundance_table_species.txt")
phyla = read_tsv("merged_abundance_table_phyla.txt")
phyla_t = phyla %>% 
  pivot_longer(!Sample:Treatment, names_to = "Phyla", values_to = "Relative abundance")
metadata = read_tsv("chioma_Sept20_metadata.txt", col_types = "cccfffddfffcf")
species_t = species[, colSums(species != 0) > 0] %>% pivot_longer(-body_site) %>% 
  pivot_wider(names_from=body_site, values_from=value)
species_meta = inner_join(metadata, species_t) %>% filter(Treatment == "Saline" | Treatment == "Bleomycin") %>% 
  filter(Floor != "Germ-free")
species_meta_absl1 = species_meta %>% 
  filter(Floor == "ABSL1")
species_meta_absl2 = species_meta %>% 
  filter(Floor == "ABSL2")
```

## Calculating distance matrices

Bray-Curtis and Jaccard distance matrices are calculated in vegan.

```{r distances}
bray = vegdist(species_meta[,14:55], method = "bray")
jaccard = vegdist(species_meta[14:55], method = "jaccard", 
                  binary = T)
bray_absl1 = vegdist(species_meta_absl1[,14:55], method = "bray")
jaccard_absl1 = vegdist(species_meta_absl1[14:55], method = "jaccard", 
                  binary = T)
bray_absl2 = vegdist(species_meta_absl2[,14:55], method = "bray")
jaccard_absl2 = vegdist(species_meta_absl2[14:55], method = "jaccard", 
                  binary = T)
```

## PERMANOVAs

A PERMANOVA is performed to assess the influence of both floor (ABSL1 vs. ABSL2) and treatment (bleomycin vs. saline). 

We will first look at composition taking species abundance into account (Bray-Curtis metric).

```{r permanova_bray}
adonis2(bray ~ Floor + Treatment, data=species_meta, 
        by = "margin", permutations = 4999)

adonis2(bray_absl1 ~ Treatment, data=species_meta_absl1, 
        by = "margin", permutations = 4999)

adonis2(bray_absl2 ~ Treatment, data=species_meta_absl2, 
        by = "margin", permutations = 4999)
```

And then at presence-absence (Jaccard metric).

```{r permanova_jaccard}
adonis2(jaccard ~ Floor + Treatment, data=species_meta, 
        by = "margin", permutations = 4999)

adonis2(jaccard_absl1 ~ Treatment, data=species_meta_absl1, 
        by = "margin", permutations = 4999)

adonis2(jaccard_absl2 ~ Treatment, data=species_meta_absl2, 
        by = "margin", permutations = 4999)
```

## Phyla bar plots

After collapsing the table at the phyla level, we constructed bar plots to visualize differences between the ABSL1 and ABSL2 floors.

```{r barplot}
tiff(file="phyla_barplot.tif", res=300, width=12, height=6.5, units="in")
phyla_plot = ggbarplot(phyla_t, x = "Sample", y = "Relative abundance",
          fill = "Phyla", color = "Phyla", palette = "Set1",
          width = 1)
ggpar(phyla_plot, legend = "right") +  
          rremove("x.text") + rremove("xlab")
dev.off()
```

## NMDS plots

We constructed a non-linear multidimensional scaling plot for the Bray-Curtis metric.

```{r nmds_bray}
mds_otus_bray<-metaMDS(bray, k=2, trymax=499)
mds_otus_bray_points<-mds_otus_bray$points
mds_otus_bray_points2<-merge(x=mds_otus_bray_points, 
                             y = species_meta, 
                             by.x = "row.names", 
                             by.y = "row.names")

mds_otus_bray_points2$Floor = factor(mds_otus_bray_points2$Floor,
                                     levels = c("ABSL1", "ABSL2",
                                                "Germ-free"))

tiff(file="nmds_plot_bray.tif", res=300, width=8, height=6, units="in")
bray_plot <- ggplot(mds_otus_bray_points2, 
                    aes(x = MDS1, y = MDS2, color = Floor, 
                        shape = Treatment)) +
  geom_point(size=3) + scale_color_brewer(palette = 'Set1') +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'), 
        legend.key=element_blank()) + 
  theme(axis.title.x=element_text(size=rel(2)), 
        axis.title.y=element_text(size=rel(2)),
        plot.title = element_text(size=rel(3)),
        legend.title = element_text(size=rel(2)),
        legend.text = element_text(size = rel(1.8))) +
  stat_ellipse(aes(x = MDS1, y = MDS2, group = Treatment, 
                   linetype = Treatment))
bray_plot
dev.off()
bray_plot
```
PDF plots are included below
```{r bray_pdf}
pdf(file="nmds_plot_bray.pdf", width=8, height=6)
bray_plot
dev.off()
```

We constructed a non-linear multidimensional scaling plot for the Jaccard metric.

```{r nmds_jaccard}
mds_otus_jaccard<-metaMDS(jaccard, k=2, trymax=499)
mds_otus_jaccard_points<-mds_otus_jaccard$points
mds_otus_jaccard_points2<-merge(x=mds_otus_jaccard_points, 
                             y = species_meta, 
                             by.x = "row.names", 
                             by.y = "row.names")

mds_otus_jaccard_points2$Floor = factor(mds_otus_jaccard_points2$Floor,
                                     levels = c("ABSL1", "ABSL2",
                                                "Germ-free"))

tiff(file="nmds_plot_jaccard.tif", res=300, width=8, height=6, units="in")
jaccard_plot <- ggplot(mds_otus_jaccard_points2, 
                    aes(x = MDS1, y = MDS2, color = Floor, 
                        shape = Treatment)) +
  geom_point(size=3) + scale_color_brewer(palette = 'Set1') +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'), 
        legend.key=element_blank()) + 
  theme(axis.title.x=element_text(size=rel(2)), 
        axis.title.y=element_text(size=rel(2)),
        plot.title = element_text(size=rel(3)),
        legend.title = element_text(size=rel(2)),
        legend.text = element_text(size = rel(1.8))) +
  stat_ellipse(aes(x = MDS1, y = MDS2, group = Treatment, 
                   linetype = Treatment))
jaccard_plot
dev.off()
jaccard_plot
```
PDF plots are included below
```{r bray_pdf}
pdf(file="nmds_plot_jaccard.pdf", width=8, height=6)
jaccard_plot
dev.off()
```

## Alpha diversity
```{r alpha}
shannond = diversity(species_meta[,14:55], index = "shannon")

shannone = shannond/log(specnumber(species_meta[,14:55]))

speciesr = specnumber(species_meta[,14:55])

diversity = cbind(species_meta[,1:14], shannond, shannone, speciesr)

diversity$Floor = factor(diversity$Floor,
                                     levels = c("ABSL1", "ABSL2"))

div = kruskal.test(shannond ~ Floor, 
                        data = diversity)
div

divt = kruskal.test(shannond ~ Treatment, 
                        data = diversity)
divt

even = kruskal.test(shannone ~ Floor, 
                        data = diversity)
even

diversity_nogerm = diversity %>% filter(Floor != "Germ-free")

div = wilcox.test(shannond ~ Floor, 
                        data = diversity_nogerm)
div

even = wilcox.test(shannone ~ Floor, 
                        data = diversity_nogerm)
even

rich = wilcox.test(speciesr ~ Floor, data = diversity_nogerm)
rich

diversity_nogerm_absl1 = diversity %>% filter(Floor != "Germ-free") %>% 
  filter(Floor == "ABSL1")

divt = wilcox.test(shannond ~ Treatment, 
                        data = diversity_nogerm_absl1)
divt

event = wilcox.test(shannone ~ Treatment, 
                        data = diversity_nogerm_absl1)
event

richt = wilcox.test(speciesr ~ Treatment, data = diversity_nogerm_absl1)
richt

diversity_nogerm_absl2 = diversity %>% filter(Floor != "Germ-free") %>% 
  filter(Floor == "ABSL2")

divt = wilcox.test(shannond ~ Treatment, 
                        data = diversity_nogerm_absl2)
divt

event = wilcox.test(shannone ~ Treatment, 
                        data = diversity_nogerm_absl2)
event

richt = wilcox.test(speciesr ~ Treatment, data = diversity_nogerm_absl2)
richt
```

### Alpha diversity graphs
```{r alpha_graphs}
div_plot = ggboxplot(diversity_nogerm, x = "Floor", 
                    y = "shannond", color = "Floor", 
                    palette = "Set1", add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Shannon Diversity") 
div_plot = ggpar(div_plot, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
div_plot

even_plot = ggboxplot(diversity_nogerm, x = "Floor", 
                    y = "shannone", color = "Floor", 
                    palette = "Set1", add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Pielou's Evenness") 
even_plot = ggpar(even_plot, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
even_plot

rich_plot = ggboxplot(diversity_nogerm, x = "Floor", 
                    y = "speciesr", color = "Floor", 
                    palette = "Set1", add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Species richness") 
rich_plot = ggpar(rich_plot, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
rich_plot

tiff(file="Floor_diversity.tif", res=300, width=12, height=3, units="in")
ggarrange(div_plot, even_plot, rich_plot, nrow = 1, ncol = 3, 
          common.legend = T, align = "h", legend = "right")
dev.off()

div_plot_t1 = ggboxplot(diversity_nogerm_absl1, x = "Treatment", 
                    y = "shannond", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Shannon Diversity") 
div_plot_t1 = ggpar(div_plot_t1, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
div_plot_t1

even_plot_t1 = ggboxplot(diversity_nogerm_absl1, x = "Treatment", 
                    y = "shannone", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Pielou's Evenness") 
even_plot_t1 = ggpar(even_plot_t1, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
even_plot_t

rich_plot_t1 = ggboxplot(diversity_nogerm_absl1, x = "Treatment", 
                    y = "speciesr", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Species richness") 
rich_plot_t1 = ggpar(rich_plot_t1, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
rich_plot_t1

div_plot_t2 = ggboxplot(diversity_nogerm_absl2, x = "Treatment", 
                    y = "shannond", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Shannon Diversity") 
div_plot_t2 = ggpar(div_plot_t2, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
div_plot_t2

even_plot_t2 = ggboxplot(diversity_nogerm_absl2, x = "Treatment", 
                    y = "shannone", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Pielou's Evenness") 
even_plot_t2 = ggpar(even_plot_t2, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
even_plot_t

rich_plot_t2 = ggboxplot(diversity_nogerm_absl2, x = "Treatment", 
                    y = "speciesr", color = "Treatment", 
                    palette = c("light gray", "dark gray"), add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Species richness") 
rich_plot_t2 = ggpar(rich_plot_t2, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title") + 
  stat_compare_means(method = "wilcox.test")
rich_plot_t2

tiff(file="Treatment_diversity.tif", res=300, width=12, height=6, units="in")
ggarrange(div_plot_t1, even_plot_t1, rich_plot_t1, 
          div_plot_t2, even_plot_t2, rich_plot_t2, 
          nrow = 2, ncol = 3, 
          common.legend = T, align = "h", legend = "right", 
          labels = c("A", "B", "C", "D", "E", "F"))
dev.off()
```

## Individual taxonomic differences

We ran a generalized linear model with a negative binomial distribution to assess the influence of treatment and floor on the relative abundance of all phyla and species detected.

```{r phyla_comp}
glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

actino = glmmTMB(Actinobacteria ~ Floor + Treatment, 
                  data = phyla, family = nbinom2)
Anova(actino)

bact = glmmTMB(Bacteroidetes ~ Floor + Treatment, 
                  data = phyla, family = nbinom2)
Anova(bact)

firm = glmmTMB(Firmicutes ~ Floor + Treatment, 
                  data = phyla, family = nbinom2)
Anova(firm)

proteo = glmmTMB(Proteobacteria ~ Floor + Treatment, 
                  data = phyla, family = nbinom2)
Anova(proteo)

verru = glmmTMB(Verrucomicrobia ~ Floor + Treatment, 
                  data = phyla, family = nbinom2)
Anova(verru)
```

```{r taxa_comp}
species_meta_comp = as.data.frame(species_meta)

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

floor_estimatematrix = mat.or.vec(42,2)
floor_pvaluematrix = mat.or.vec(42,2)
treatment_estimatematrix = mat.or.vec(42,2)
treatment_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_comp[,i]
  b = try(glmmTMB(variable ~ Floor + Treatment, 
                  data = species_meta_comp, family = nbinom2))
  anova = Anova(b)
  floor_estimatematrix[i-13,2] = anova[1,1]
  floor_pvaluematrix[i-13,2] = anova[1,3]
  treatment_estimatematrix[i-13,2] = anova[2,1]
  treatment_pvaluematrix[i-13,2] = anova[2,3]
  floor_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  floor_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
}

floor = bind_cols(as.data.frame(floor_estimatematrix[,1:2]), 
                     as.data.frame(floor_pvaluematrix[,2]))
write.csv(floor, "floor.csv")
floor_nona = filter(floor, floor[,3] != "NaN") 
floor_nona[,3] = as.numeric(as.character(floor_nona[,3]))
floor_corrected = bind_cols(floor_nona, 
                               as.data.frame(fdrtool(floor_nona[,3], 
                                                     statistic = "pvalue", plot = F)))
write.csv(floor_corrected, "floor_corrected.csv")

treatment = bind_cols(as.data.frame(treatment_estimatematrix[,1:2]), 
                      as.data.frame(treatment_pvaluematrix[,2]))
write.csv(treatment, "treatment.csv")
treatment_nona = filter(treatment, treatment[,3] != "NaN") 
treatment_nona[,3] = as.numeric(as.character(treatment_nona[,3]))
treatment_corrected = bind_cols(treatment_nona, 
                                as.data.frame(fdrtool(treatment_nona[,3], 
                                                      statistic = "pvalue", plot = F)))
write.csv(treatment_corrected, "treatment_corrected.csv")

a = glmmTMB(Parasutterella_excrementihominis ~ Floor + Treatment, 
            data = species_meta_comp, family = nbinom2)
summary(a)
Anova(a)
summary(glht(a,linfct=mcp(Floor="Tukey")))
```

```{r taxa_comp_nog}
species_meta_comp = as.data.frame(species_meta) %>% 
  filter(Floor != "Germ-free")

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}

floor_estimatematrix = mat.or.vec(42,2)
floor_pvaluematrix = mat.or.vec(42,2)
treatment_estimatematrix = mat.or.vec(42,2)
treatment_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_comp[,i]
  b = try(glmmTMB(variable ~ Floor + Treatment, 
                  data = species_meta_comp, family = nbinom2))
  anova = Anova(b)
  floor_estimatematrix[i-13,2] = anova[1,1]
  floor_pvaluematrix[i-13,2] = anova[1,3]
  treatment_estimatematrix[i-13,2] = anova[2,1]
  treatment_pvaluematrix[i-13,2] = anova[2,3]
  floor_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  floor_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_estimatematrix[i-13,1] = names(species_meta_comp)[i]
  treatment_pvaluematrix[i-13,1] = names(species_meta_comp)[i]
}

floor = bind_cols(as.data.frame(floor_estimatematrix[,1:2]), 
                     as.data.frame(floor_pvaluematrix[,2]))
write.csv(floor, "floor_nogerm.csv")
floor_nona = filter(floor, floor[,3] != "NaN") 
floor_nona[,3] = as.numeric(as.character(floor_nona[,3]))
floor_corrected = bind_cols(floor_nona, 
                               as.data.frame(fdrtool(floor_nona[,3], 
                                                     statistic = "pvalue", plot = F)))
write.csv(floor_corrected, "floor_nogerm_corrected.csv")

treatment = bind_cols(as.data.frame(treatment_estimatematrix[,1:2]), 
                      as.data.frame(treatment_pvaluematrix[,2]))
write.csv(treatment, "treatment_nogerm.csv")
treatment_nona = filter(treatment, treatment[,3] != "NaN") 
treatment_nona[,3] = as.numeric(as.character(treatment_nona[,3]))
treatment_corrected = bind_cols(treatment_nona, 
                                as.data.frame(fdrtool(treatment_nona[,3], 
                                                      statistic = "pvalue", plot = F)))
write.csv(treatment_corrected, "treatment_nogerm_corrected.csv")

a = glmmTMB(Parasutterella_excrementihominis ~ Floor + Treatment, 
            data = species_meta_comp, family = nbinom2)
summary(a)
Anova(a)
summary(glht(a,linfct=mcp(Floor="Tukey")))
```
## Single Taxa Graphs

One taxa, Parasutterella excrementihominis, was more abundant in mice on the non-infectious floor compared with both the infectious and germ-free floors.

```{r graphs}
species_meta_comp$Floor = factor(species_meta_comp$Floor,
                                     levels = c("ABSL1", "ABSL2",
                                                "Germ-free"))

para = ggboxplot(species_meta_comp, x = "Floor", 
                    y = "Parasutterella_excrementihominis", color = "Floor", 
                    palette = "Set1", add = "jitter", 
                    add.params = list(fill = "white"), 
                 ylab = "Relative Abundance", title = "Parasutterella excrementihominis") 
para = ggpar(para, legend = "right") + 
  rremove("xlab") + rremove("x.text") + 
  rremove("legend.title")
para

tiff(file="Floor_taxa.tif", res=300, width=4, height=3, units="in")
para
dev.off()
```

And PDF output
```{r graphs_pdf}
pdf("Floor_taxa.pdf", width = 4, height = 3)
para
dev.off()
```

#Integrating physiological data with taxonomic data

In order to identify if the relative abundance of any taxa were associated with weight loss, gp130 content, or immune markers.

```{r import_physio}
phys_data = read_csv("physiology_data.csv")
species_meta_phys = species_meta %>% 
  left_join(phys_data, by = "#sampleid")
species_meta_phys2 = as.data.frame(species_meta_phys)
```

We ran Spearman correlations to assess the influence of physiological variables on the relative abundance of all species detected.

```{r taxa_weight}
weight_estimatematrix = mat.or.vec(42,2)
weight_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`Weight Loss`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  weight_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  weight_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  weight_estimatematrix[i-13,2] = e[[1]]
  weight_pvaluematrix[i-13,2] = p
}

weight = bind_cols(as.data.frame(weight_estimatematrix[,1:2]), 
                     as.data.frame(weight_pvaluematrix[,2]))
write.csv(weight, "weight_loss.csv")
weight_nona = filter(weight, weight[,3] != "NaN") 
weight_nona[,3] = as.numeric(as.character(weight_nona[,3]))
weight_corrected = bind_cols(weight_nona, 
                               as.data.frame(fdrtool(weight_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(weight_corrected, "weight_loss_corrected.csv")
```

And then collagen content

```{r taxa_collagen}
collagen_estimatematrix = mat.or.vec(42,2)
collagen_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`Collagen Content`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  collagen_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  collagen_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  collagen_estimatematrix[i-13,2] = e[[1]]
  collagen_pvaluematrix[i-13,2] = p
}

collagen = bind_cols(as.data.frame(collagen_estimatematrix[,1:2]), 
                     as.data.frame(collagen_pvaluematrix[,2]))
write.csv(collagen, "collagen_loss.csv")
collagen_nona = filter(collagen, collagen[,3] != "NaN") 
collagen_nona[,3] = as.numeric(as.character(collagen_nona[,3]))
collagen_corrected = bind_cols(collagen_nona, 
                               as.data.frame(fdrtool(collagen_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(collagen_corrected, "collagen_loss_corrected.csv")
```
CD4+IL-6+T cells
```{r taxa_il6}
il6_estimatematrix = mat.or.vec(42,2)
il6_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`CD4+IL-6+ T cells`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  il6_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  il6_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  il6_estimatematrix[i-13,2] = e[[1]]
  il6_pvaluematrix[i-13,2] = p
}

il6 = bind_cols(as.data.frame(il6_estimatematrix[,1:2]), 
                     as.data.frame(il6_pvaluematrix[,2]))
write.csv(il6, "il6_loss.csv")
il6_nona = filter(il6, il6[,3] != "NaN") 
il6_nona[,3] = as.numeric(as.character(il6_nona[,3]))
il6_corrected = bind_cols(il6_nona, 
                               as.data.frame(fdrtool(il6_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(il6_corrected, "il6_loss_corrected.csv")
```
CD4+PD-1+T cells
```{r taxa_pd1}
pd1_estimatematrix = mat.or.vec(42,2)
pd1_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`CD4+PD-1+ T cells`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  pd1_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  pd1_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  pd1_estimatematrix[i-13,2] = e[[1]]
  pd1_pvaluematrix[i-13,2] = p
}

pd1 = bind_cols(as.data.frame(pd1_estimatematrix[,1:2]), 
                     as.data.frame(pd1_pvaluematrix[,2]))
write.csv(pd1, "pd1_loss.csv")
pd1_nona = filter(pd1, pd1[,3] != "NaN") 
pd1_nona[,3] = as.numeric(as.character(pd1_nona[,3]))
pd1_corrected = bind_cols(pd1_nona, 
                               as.data.frame(fdrtool(pd1_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(pd1_corrected, "pd1_loss_corrected.csv")
```

CD4+PD-1+IL-17A+T cells

```{r taxa_il17}
il17_estimatematrix = mat.or.vec(42,2)
il17_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`CD4+PD-1+IL-17A+ T cells`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  il17_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  il17_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  il17_estimatematrix[i-13,2] = e[[1]]
  il17_pvaluematrix[i-13,2] = p
}

il17 = bind_cols(as.data.frame(il17_estimatematrix[,1:2]), 
                     as.data.frame(il17_pvaluematrix[,2]))
write.csv(il17, "il17_loss.csv")
il17_nona = filter(il17, il17[,3] != "NaN") 
il17_nona[,3] = as.numeric(as.character(il17_nona[,3]))
il17_corrected = bind_cols(il17_nona, 
                               as.data.frame(fdrtool(il17_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(il17_corrected, "il17_loss_corrected.csv")
```
CD4+pSTAT3+T cells
```{r taxa_pstat3}
pstat3_estimatematrix = mat.or.vec(42,2)
pstat3_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`CD4+pSTAT3+ T cells`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  pstat3_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  pstat3_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  pstat3_estimatematrix[i-13,2] = e[[1]]
  pstat3_pvaluematrix[i-13,2] = p
}

pstat3 = bind_cols(as.data.frame(pstat3_estimatematrix[,1:2]), 
                     as.data.frame(pstat3_pvaluematrix[,2]))
write.csv(pstat3, "pstat3_loss.csv")
pstat3_nona = filter(pstat3, pstat3[,3] != "NaN") 
pstat3_nona[,3] = as.numeric(as.character(pstat3_nona[,3]))
pstat3_corrected = bind_cols(pstat3_nona, 
                               as.data.frame(fdrtool(pstat3_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(pstat3_corrected, "pstat3_loss_corrected.csv")
```
CD4+gp130+T cells
```{r taxa_gp130}
gp130_estimatematrix = mat.or.vec(42,2)
gp130_pvaluematrix = mat.or.vec(42,2)

for(i in 14:55) {
  variable = species_meta_phys2[,i]
  a = cor.test(x = variable, 
               y = species_meta_phys2$`CD4+gp130+ T cells`, 
                  data = species_meta_phys2,
               method = "spearman",
               na.action = na.omit)
  p = a$'p.value'
  e = a$'estimate'
  gp130_estimatematrix[i-13,1] = names(species_meta_phys2)[i]
  gp130_pvaluematrix[i-13,1] = names(species_meta_phys2)[i]
  gp130_estimatematrix[i-13,2] = e[[1]]
  gp130_pvaluematrix[i-13,2] = p
}

gp130 = bind_cols(as.data.frame(gp130_estimatematrix[,1:2]), 
                     as.data.frame(gp130_pvaluematrix[,2]))
write.csv(gp130, "gp130_loss.csv")
gp130_nona = filter(gp130, gp130[,3] != "NaN") 
gp130_nona[,3] = as.numeric(as.character(gp130_nona[,3]))
gp130_corrected = bind_cols(gp130_nona, 
                               as.data.frame(fdrtool(gp130_nona[,3], statistic = "pvalue", 
                                                     plot = F)))
write.csv(gp130_corrected, "gp130_loss_corrected.csv")
```
